Beyond Adaptive Web Design

### 2014-03-15 paper/paper.tex
fix: fix typos

The power of the web is its ubiquity. No one knows what the landscape is going to appear in a couple years, but there's a reasoneable probability the new devices in a few years from now will have access to the web. Because of this, we need to preserve and embrace the web's ubiquity. This requires us to provide full web experiences regardless of how and where people access the web. It's important to remain in creating flexible interfaces that can adapt to any screen size. Performance is often less important to everything else. ``Progressive enhancement'', feature detection and many other techniques allow us to increase experience that allow us to support more devices while still optimizing for the finest. The key aspect of ``Future Friendly'' thinking is to acknowledge and embrace unpredictability. Nobody knows where things are going but the one thing we can trust is the evolution.

@@

Touch is no longer just isolated to smartphones and tablets. Many big sreens in laptops or desktops nowadays are touch capable. Every desktop design has to be touch-friendly now.

@@

The speech input aims to provide an alternative input method for web applications, without using a keyboard or other physical device. This is no longer science fiction but reality. Many companies rely on it and deploy it to the modern devices and software. It could be also benefitial for older or disabled users~\cite{SpeechRecognition} and can easily become the next big thing.

@@

So what we should do about it? We can create reuseable components and conditionaly load resources. Web components allow us to use custom HTML elements in the browser. We can enclose a simple element or an entire application logic within an HTML element~\cite{webcomponents}. We can build Web applications in a reusable, commutable and encapuslated way.

@@

We have detected several web componenets and input methods and tried to verify which of them provide better experience for the website users and are also useful for the web developers. There have been made series of experiments on both adaptive input methods and adaptive web components. To prove the usefulness of the proposed concept we have designed and implemented several reusable modules and components and we have used them in propotype web projects. All of them have been released as open source and are available on GitHub~\footnotemark.

@@

Speech input provides a great opportunity to take a complete control over web application using only voice commands. For this purpose we have used web browser's Speech API for continuous speech recognition and built a configurable and reuseable module. Speech recognition is however currently possible online only.

Using speech recognition we get an utterance string which can be compared to voice commands from our configuration. We use regular expressions for this purpose with attached callback methods. These configurations are bound to utterance watchers and the callback is called when voice command is recognised. Regular expressions are benefitial since we can set up a single watcher for dynamical commands where one command can be used for multiple subjects.

But there are many more interesting things we can achieve using speech input. We can even use multiple languages for speech recognition. One of the biggest problems of the voice commands is however incorrect speech recognition. To solve this issue we can conditionalize regular expressions or use uterance error correction.

@@

We have access to 3 axis rotation informations which are recalculated every time on device orientation change. We remember starting device orientation and compare it to the current one. Based on this substraction we can make a decision which direction we scroll and how fast. The bigger the orientation difference is, the faster we scroll.

@@

We use the Mobile First principe so we make a default elemet static and pull in a static image. This approach saves a lot of web traffic and becasue of limited mobile internet plans also user's money. We can then detect when it's appropriate to download the embedded map or to launch the native mobile application.

@@

Adaptive web components are a great to save an unwonted web traffic. They also icrease user experience because the web page produces less requests and loads faster. However the limitations also exist because many websites need custom elements and design. As a result the adaptive web components produce less interest than adaptive input methods, but are the necessary part of the Web.

>>>>

The power of the web is its ubiquity. No one knows what the landscape is going to appear in a couple years, but there's a reasonable probability the new devices in a few years from now will have access to the web. Because of this, we need to preserve and embrace the web's ubiquity. This requires us to provide full web experiences regardless of how and where people access the web. It's important to remain in creating flexible interfaces that can adapt to any screen size. Performance is often less important to everything else. ``Progressive enhancement'', feature detection and many other techniques allow us to increase experience that allow us to support more devices while still optimizing for the finest. The key aspect of ``Future Friendly'' thinking is to acknowledge and embrace unpredictability. Nobody knows where things are going but the one thing we can trust is the evolution.

@@

Touch is no longer just isolated to smartphones and tablets. Many big screens in laptops or desktops nowadays are touch capable. Every desktop design has to be touch-friendly now.

@@

The speech input aims to provide an alternative input method for web applications, without using a keyboard or other physical device. This is no longer science fiction but reality. Many companies rely on it and deploy it to the modern devices and software. It could be also beneficial for older or disabled users~\cite{SpeechRecognition} and can easily become the next big thing.

@@

So what we should do about it? We can create reusable components and conditionally load resources. Web components allow us to use custom HTML elements in the browser. We can enclose a simple element or an entire application logic within an HTML element~\cite{webcomponents}. We can build Web applications in a reusable, commutable and encapuslated way.

@@

We have detected several web components and input methods and tried to verify which of them provide better experience for the website users and are also useful for the web developers. There have been made series of experiments on both adaptive input methods and adaptive web components. To prove the usefulness of the proposed concept we have designed and implemented several reusable modules and components and we have used them in propotype web projects. All of them have been released as open source and are available on GitHub~\footnotemark.

@@

Speech input provides a great opportunity to take a complete control over web application using only voice commands. For this purpose we have used web browser's Speech API for continuous speech recognition and built a configurable and reusable module. Speech recognition is however currently possible online only.

Using speech recognition we get an utterance string which can be compared to voice commands from our configuration. We use regular expressions for this purpose with attached callback methods. These configurations are bound to utterance watchers and the callback is called when voice command is recognised. Regular expressions are beneficial since we can set up a single watcher for dynamical commands where one command can be used for multiple subjects.

But there are many more interesting things we can achieve using speech input. We can even use multiple languages for speech recognition. One of the biggest problems of the voice commands is however incorrect speech recognition. To solve this issue we can conditionalize regular expressions or use utterance error correction.

@@

We have access to 3 axis rotation informations which are recalculated every time on device orientation change. We remember starting device orientation and compare it to the current one. Based on this subtraction we can make a decision which direction we scroll and how fast. The bigger the orientation difference is, the faster we scroll.

@@

We use the Mobile First principe so we make a default element static and pull in a static image. This approach saves a lot of web traffic and becasue of limited mobile internet plans also user's money. We can then detect when it's appropriate to download the embedded map or to launch the native mobile application.

@@

Adaptive web components are a great to save an unwonted web traffic. They also increase user experience because the web page produces less requests and loads faster. However the limitations also exist because many websites need custom elements and design. As a result the adaptive web components produce less interest than adaptive input methods, but are the necessary part of the Web.
